Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-10-08 methods_paper/fabl-paper.tex
added computational complexity

\bigskip
\begin{abstract}
In this paper, we propose the \emph{fast beta linkage} (\texttt{fabl}) methods, which extends a recent Bayesian Fellegi Sunter model to allow for parallel computing. In addition, we propose using hashing techniques to hasten calculations and to reduce the overall computational complexity. To achieve a scalable model, we propose independent priors in order to create computational speeds and gains, where we relax the one-to-one matching assumption. Second, we resolve the one-to-one matching issue via a post processing step. Third, we derive conditional distributions, a Gibbs sampler, and corresponding Bayes estimates for our proposed model. Fourth, we explore other computational speeds up based upon storage efficient indexing. Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices, where we provide comparisons to the \text{BRL} package in CRAN.

\vspace{1ex}
{
\color{blue}
In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space. This allows us to employ hashing techniques that hasten calculations and reduce computational complexity, complete pairwise record comparisons over large datasets through parallel computing, and reduce memory costs through a new technique called \emph{storage efficient indexing}. Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.

\vspace{1ex}
QUESTION: Does this abstract necessarily need to read like an outline?
}
\end{abstract}

@@

\textbf{Our Contribution} 
In this paper, we propose a Bayesian extension of \cite{sadinle_bayesian_2017} (\text{BRL}) through a modification to the model specification that allows for parallel computing of the linkage parameter, using hashing techniques to hasten calculations, and to reduce computational complexity.  Specifically, \cite{sadinle_bayesian_2017} proposes a prior distribution that strictly enforces one-to-one matching. Instead, we propose independent priors (breaking the one-to-one matching) in order to create computational speeds and gains. Second, we resolve the one-to-one matching issue via a post processing step. Third, we derive conditional distributions, a Gibbs sampler, and corresponding Bayes estimates for our proposed model. Fourth, we explore other computational speeds up based upon storage efficient indexing. Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices. 
To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl},
pronounced ``fable'')

@@

In this section, we deviate from the approach of \cite{sadinle_bayesian_2017}, and propose a more scalable prior on the matching labelings $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2}),$ where recall $Z_j \in \{1,2, 3, n_1, n_1 + j\}.$

While the rest of the model has been proposed previously, to our knowledge, this particular prior has not been proposed before.

\textcolor{red}{Brian: could you please lay out your proposal here more clear regarding the notation above and add as much details as possible regarding your approach. More detail is better than less. It will be important to contrast this with the prior that Sadinle uses and note that you're dropping the constraint. (Make sure to include any proofs/derivations that might not be obvious to a reader given this is a major contribution to the paper).}

@@

{
	\color{blue}
	Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\lambda$, and that each record $i \in \bm{X}_1$ is equally likely to be that match. $\lambda$ itself is given a prior distribution to reflect prior beliefs about the overall rate of matchingness in the data. One choice of uninformative prior might be \(\lambda \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\lambda \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).
}

@@

{
\color{blue}
There is a clear relationship between our proposed model and that of \cite{sadinle_bayesian_2017}. 
\cite{sadinle_bayesian_2017} constructs a prior distribution on the entire \(\mathbf{Z}\) vector, called the \emph{beta distribution for bipartite matching}, given by
$$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior induces a Gibbs sampler that strictly enforces one-to-one matching,  previously matched records from the set of candidate
records when sampling \(Z_j\). This creates a dependency that makes the
sampler \emph{inherently serial}.

On the other hand, we propose independent priors for each \(Z_j\), creating a sampler that is \emph{embarassingly parallel}, allowing for significant computational gains. More importantly, since only the agreement pattern of \(Z_j\) is used for
calculations within the Gibbs sampler, and not the particular record
label, this sampling is only needed at the level of the unique agreement patterns, which offers additional computational savings.  In doing so however, we thereby weaken the one-to-one requirement from
\texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be
matched to at most one record in \(\bm{X}_1\), but allows for the possibility
that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). \textcolor{blue}{We resolve contradictions when calcuating the Bayes estimate, which is explained in Section 4.5}

NOTE TO BEKA: I think it will be best to do the "post-processing step" along with the Bayes Estimate. Its equivalent to minimizing the posterior loss subject to the constraint that $Z_j \neq Z_{j'}, \forall j \neq j'$.
}

@@

{
\color{blue}
Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$. It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of "many to one" matches indicate that the assumption of no duplicates within files may not be reasonable. We will explore this issue more deeply through the El Salvador homicide case study.

To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$. In the event that two matches. In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, this constraint means we accept the match with the highest posterior probability, and declare the other to have no match.  A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities, and is justified within this Bayesian framework.
}

@@

\textcolor{red}{Brian: how large can we go and how fast can we go?}

\section{Efficient Posterior Inference}\label{sec:efficient-posterior}

@@

\begin{algorithm}
	\caption{Summary of fabl algorithm}
	\begin{algorithmic}[1]

@@

\section{Simulation Studies}
\label{sec:simulations}
In this section, we consider a simulation study, nearly identical to that in \cite{sadinle2017}. \textcolor{red}{Brian: can you explain how it's the same versus different than what it is given in Sadinle as this would be the most helpful. Do you utilize Mauricio's exact files and approach and then do something else? If so, I would state this for clarity.} 
{
\color{blue}For Beka: I have re-run the simulation exactly as Sadinle implemented it. Here is the rewritten paragraph.

\vspace{1em}
In this section, we replicate a simulation study from \cite{sadinle2017} to compare \texttt{fabl} against \texttt{BRL} on several simulated datasets with varying amounts of error and overlap. We use first name, last name, age, and occupation for this linkage, and create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package. See Table~\ref{Tab:sadinle_simulation_cutoffs} for details on the construction of comparison vectors. Each simulation identifies duplicated individuals between two datasets, each with 500 records. We conduct linkage when matching records exhibit 1, 2, and 3 errors across the four fields, and when there are 50, 250, and 450 individuals in common across datasets. We use flat priors for all $m$ and $u$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.
\vspace{1em}
}

\begin{table}
	\centering
	\begin{tabular}[t]{ll|llll}
		\hline
		\multicolumn{2}{c|}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
		\cline{3-6}
		Fields & Similarity & 1 & 2 & 3 & 4\\
		\hline
		First and Last Name & Levenstein & 0 & (0, .25] & (.25, .5] & (5, .1]\\
		\hline
		Age and Occupation & Binary & Agree & Disagree &  & \\
		\hline
	\end{tabular}
\caption{Construction of comparison vectors for accuracy study with simulated datasets}
\label{Tab:sadinle_simulation_cutoffs}
\end{table}

\textcolor{red}{Brian: If you look in Sadinle (2017), did you use Table 2 or something else? Can you construct something similar to Figure 3 and Figure 4 as it's likely a reviewer will look at this paper and ask why you didn't do this exact comparison.}

{
\color{blue}For Beka: Yes, I used Table 2 (those are default settings in BRL). See above comment

My Figure 1 is equivalent to Sadinle's figure 3. He compares against Fellegi Sunter, whereas I compare against BRL. I found his style of graphic difficult to read, especially because I wanted to show F-measure in addition to recall and precision.

Sadinle's Figure 4 just shows the results when you add the option of not declaring matching status (leaving it undecided by the model). This didn't seem to be so relevant in my paper, so I didn't have it before. I just completed the simulation, and I'll include it under a seperate section.
}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot2} 
\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017. For each level of overlap and each level of error, we have 100 paired sets of 500 records.}
\label{fig:sadinle_simulation}
\end{center}
\end{figure}

@@

{
\color{blue}
Lastly, we review the performance of \texttt{fabl} and \texttt{BRL} using loss functions that allow partial estimates of the linkage parameter. By leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match. In this context, we are no longer focused on finding all true matches, but rather protecting against false matches. Thus, instead of recall, we use the \emph{negative predictive value} (NPV), defined as the proportion of non-links that are actual non-matches. Mathematically, $NPV = \sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j = n_1 + j)$/$\sum_{j=1}^{n_2} I(\hat{Z}_j = n_1 + j)$. We continue to use the precision, which is renamed the \emph{positive predictive value} in this context. Lastly, we also report that rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)$.

@@

NOTE: Sadinle only shows results for the 10\% overlap case. Should I inlcude all of this?
NOTE: When I ran this, I didn't realize I needed to calculate Rejection Rate. I'll have to run the simulation again to get this. 
}

\begin{figure}[htbp]
\begin{center}
	\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial} 
	\caption{Negative predictive value (NPV) and positive predictive value (PPV) on simulated datasets}
	\label{fig:sadinle_simulation_partial}
\end{center}
\end{figure}

@@

\textcolor{red}{Brian: I have some general questions here. Are you using blocking for BRL to scale? Could you use the hashing proposal for BRL. It's important that the scale comparison be as fair as possible or there will be some pushback here. My suggestion around all this is to look at the computational complexity of both methods as it's hard to attack this.}

{
\color{blue} For Beka: I'm not using any blocking for either method.

Some of the hashing techniques could work for BRL (updating the m and u parameters), and they technically could be used for updating Z, but it would be much less useful because BRL requires knowing which exact record you match to, not just the agreement pattern for that record. Its not feasible for me try making those changes since its coded in C.

I'll work out the computational complexity. That can be one of the things we nail down on Thursday!}

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rrr}
		\hline
		& m & u \\ 
		\hline
		fname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
		lname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
		day & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{30}, \frac{29}{30}\right)$ \\ 
		month & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
		year & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
		\hline
	\end{tabular}
\caption{Distributions used for $m$ and $u$ probabilties in simulation studies}\label{Tab:distributions}
\end{table}

\begin{figure}[ht]

@@

\begin{figure}[t]

@@

% Assuming a 16GB personal machine, standard record linkage methods using all-to-all comparisons can only function on tasks where the matrix of comparison vectors $\Gamma$ is less than 8GB. In the context of the above simulation, each comparison vector contains four integers, collectively requiring 16 bytes of memory; assuming files of equal lengths, this means the largest linkage task possible is around $7000 \times 7000$.

@@

Specifically, we demonstrate how partitioning and hashing the data through \texttt{fabl} method allows us undertake significantly larger linkage tasks. We concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{} to create two larger files, each of 20,000 records (40,000 records total). Working on one machine, we sequentially compare records across chunks, hash results, and then synthesize summary statistics for the entire simulation. Under standard Fellegi Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting a final comparison matrix about 6.4 GB in size.

For simplicity, we partition one dataset into 20 smaller chunks, and leave the second dataset fully intact. We compare records, hash results, and then synthesize summary statistics for all 20 chunk comparisons. The resulting data object is now only 90 MB, about 1\% the size of the object required under the standard method. Executed sequentially, these comparisons and the Gibbs sampler take about one hour to run; using distributed computing. However, this could be sped up.

Turning to evaluation metrics, this simulation achieved 96.5\% recall and 97.7\% precision, with an overall F-measure of 97.1\% F-measure. The reader will note that this slightly worse performance than witnessed in the smaller simulation studies; this is expected because it is naturally more difficult to link more files with the same amount of information. With more linkage fields \texttt{fabl} maintains the high accuracy seen above. \textcolor{red}{Brian: you should look to see if you fix these in the post-processing step as this would be interesting/important.}

\textcolor{blue}{This is with post processing already. All I meant here was that is that small tasks require less information. If you do a $500 \times 500$ linkage task and you only have first and last name (for example), thats likely enough to get all the correct matches. However, once you do $20000 \times 20000$, its more likely that more than one person in the second file has the same first and last name, so you need more information to distinguish. I mentioned before that the simulated files actually have 6 fields of information, but I only used 4 originally because thats what Sadinle did. I can use all 6 though to see what happens?}

 \textcolor{red}{How long does it take on the same machine to run BRL. Can you run this? This can close this section out!}

 \textcolor{blue}{When I use the compareRecords from BRL to do the pairwise comparison on the concatenated files, its too  big for R to handle! One could do hashing, reduce the data down to summary statistics, and alter the BRL to use that new object, but that would be a whole different project!}

\section{Application to Civilian Casualties from the El Salvadoran Civil War}
\label{sex:case-studies}

@@

The country of El Salvador was immersed in civil war from 1980 to 1991,
and throughout the time, several organizations attempted to document
casualties of the conflict. When estimating the total number of
casualties, one cannot simply sum the numbers recorded by each
organization, as it is likely that the same individuals are recorded in
multiple casualty lists. Thanks to the Human Rights Data Analysis Group (HRDAG), we utilize 
two different sources, namely,  El Rescate - Tutela
Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
acronym in Spanish). The ERTL dataset consists of digitized reports that
had been published throughout the conflict. The CDHES dataset consists
of casualties that had been reported directly to the organization, and
later digitized.

\textcolor{red}{Beka: These data sets need more information regarding how the data sets were collected as this is quite important. See Sadinle (2017) and references for info). }

@@

In this paper, we utilize records that have non-missing entries for given and last name, which results 
in \(n_1 = 4420\) files in CHDES and \(n_2 = 1323\) files
in ERTL. We standardized names to account for common misspellings
in the Spanish language. The string distance used is a modified Levenstein
distance to account for the fact that second names are often omitted.
Place of death is recorded by municipality and department within that
municipality; however, since department was missing in 95\% of records
in CHDES and 80\% of records in ERTL, this was excluded from our analyses. 
Thus, we conduct record linkage using given name, last name,
municipality, and day, month, and year of death. We again use flat
priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters. To our knowledge, we have followed the same guidelines/settings in \cite{sadinle_bayesian_2017}.
\begin{table}
\begin{tabular}[t]{ll|llll}
	\hline
	\multicolumn{2}{c|}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
	\cline{3-6}
	Fields & Similarity & 1 & 2 & 3 & 4\\
	\hline
	First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, .5] & (5, 1]\\
	\hline
	Year of Death & Absolute Difference & 0 & 1 & 2 & 3+\\
	\hline
	Month of Death & Absolute Difference & 0 & 1 & 2-3 & 4+\\
	\hline
	Day of Death & Absolute Difference & 0 & 1-2 & 3-7 & 8+\\
	\hline
	Municipality & Binary & Agree & Disagree &  & \\
	\hline
\end{tabular}
\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \citep{sadinle2018bayesian}. This set up leads to 2048 possible agreement patterns in total}\label{Tab:el_salvador_cutoffs_1}
\end{table}

>>>>

\bigskip
\begin{abstract}
%In this paper, we propose the \emph{fast beta linkage} (\texttt{fabl}) methods, which extends a recent Bayesian Fellegi Sunter model to allow for parallel computing. In addition, we propose using hashing techniques to hasten calculations and to reduce the overall computational complexity. To achieve a scalable model, we propose independent priors in order to create computational speeds and gains, where we relax the one-to-one matching assumption. Second, we resolve the one-to-one matching issue via a post processing step. Third, we derive conditional distributions, a Gibbs sampler, and corresponding Bayes estimates for our proposed model. Fourth, we explore other computational speeds up based upon storage efficient indexing. Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices, where we provide comparisons to the \text{BRL} package in CRAN.

{
	\color{blue} Thanks Beka for writing that first abstract. I think this one captures the idea better, even if its not exactly in "outline" form.
	\vspace{1em}
}

In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space. This allows us to employ hashing techniques that hasten calculations and reduce computational complexity, complete pairwise record comparisons over large datasets through parallel computing, and reduce memory costs through a new technique called \emph{storage efficient indexing}. Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.

\end{abstract}

@@

\textbf{Our Contribution} 
In this paper, we propose a Bayesian extension of \cite{sadinle_bayesian_2017} (\texttt{BRL}) through a modification to the model specification that allows for parallel computing of the linkage parameter, using hashing techniques to hasten calculations, and to reduce computational complexity.  Specifically, \cite{sadinle_bayesian_2017} proposes a prior distribution that strictly enforces one-to-one matching. Instead, we propose independent priors (breaking the one-to-one matching) in order to create computational speeds and gains. Second, we resolve the one-to-one matching issue via a post processing step. Third, we derive conditional distributions, a Gibbs sampler, and corresponding Bayes estimates for our proposed model. Fourth, we explore other computational speeds up based upon storage efficient indexing. Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices. 
To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl},
pronounced ``fable'')

@@

In this section, we deviate from the approach of \cite{sadinle_bayesian_2017}, and propose a more scalable prior on the matching labelings $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2}),$ where recall $Z_j \in \{1,2, 3, n_1, n_1 + j\}.$ While the rest of the model has been proposed previously, to our knowledge, this particular prior has not been proposed before.

@@

Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\lambda$, and that each record $i \in \bm{X}_1$ is equally likely to be that match. $\lambda$ itself is given a prior distribution to reflect prior beliefs about the overall rate of matchingness in the data. One choice of uninformative prior might be \(\lambda \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\lambda \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).

@@

There is a clear relationship between our proposed model and that of \cite{sadinle_bayesian_2017}. 
\cite{sadinle_bayesian_2017} constructs a prior distribution on the entire \(\mathbf{Z}\) vector, called the \emph{beta distribution for bipartite matching}, given by
$$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior induces a Gibbs sampler that strictly enforces one-to-one matching,  previously matched records from the set of candidate
records when sampling \(Z_j\). This creates a dependency that makes the
sampler \emph{inherently serial}.

On the other hand, we propose independent priors for each \(Z_j\), creating a sampler that is \emph{embarassingly parallel}, allowing for significant computational gains. More importantly, since only the agreement pattern of \(Z_j\) is used for
calculations within the Gibbs sampler, and not the particular record
label, this sampling is only needed at the level of the unique agreement patterns, which offers additional computational savings.  In doing so however, we thereby weaken the one-to-one requirement from
\texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be
matched to at most one record in \(\bm{X}_1\), but allows for the possibility
that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\).

@@

Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$. It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of "many to one" matches indicate that the assumption of no duplicates within files may not be reasonable. We will explore this issue more deeply through the El Salvador homicide case study.

To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$. In the event that two matches. In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, this constraint means we accept the match with the highest posterior probability, and declare the other to have no match.  A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities, and is justified within this Bayesian framework.

@@

\hypertarget{efficient-posterior}{%
	\subsection{Efficient Posterior Inference }\label{efficient-posterior}}

@@

\begin{algorithm}[h!]
	\caption{Summary of fabl algorithm}
	\begin{algorithmic}[1]

@@

\hypertarget{computational-complexity}{%
	\subsection{Computational Complexity}\label{computational-complexity}}

{
\color{blue}
The computational complexity of \texttt{fabl} is determined by its two steps: constructing comparison vectors and Gibbs sampling. The computational complexity of all pairwise comparisons across files $\mathbf{X}_1$ and $\mathbf{X}_2$ is $O(F n_1 n_2)$. \textcolor{blue}{TALK TO BEKA. I CAN DO THE SAME TRICKS AS TED HERE TO REDUCE COMPLEXITY}. Since \texttt{fabl} is able to compute these comparisons in parallel, we can split these computations across $B$ equally sized partitions, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.

Without hashing, the computational complexity of updating the $\mathbf{m}$ and $\mathbf{u}$ parameters is $O(F n_1 n_2)$. However, by doing calculations over the agreement patterns rather than the individual record, hashing reduces the complexity to $O(FP)$.

The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$. With hashing, we split this sampling into two steps; first we sample that agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record that exhibits that pattern with complexity $O(n_2)$. Thus we have reduced the complexity of the Gibbs sampler under \texttt{BRL} from $O(F n_1 n_2)$ to $O(n_2 P)$.

Overall, the computational complexity under the complete \texttt{fabl} framework is $O(\frac{F}{B} n_1 n_2) + O(n_2 P)$.
}

\section{Simulation Studies}
\label{sec:simulations}

In this section, we replicate a simulation study from \cite{sadinle2017} to compare \texttt{fabl} against \texttt{BRL} on several simulated datasets with varying amounts of error and overlap. We use first name, last name, age, and occupation for this linkage, and create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package. See Table~\ref{Tab:sadinle_simulation_cutoffs} for details on the construction of comparison vectors. Each simulation identifies duplicated individuals between two datasets, each with 500 records. We conduct linkage when matching records exhibit 1, 2, and 3 errors across the four fields, and when there are 50, 250, and 450 individuals in common across datasets. We use flat priors for all $m$ and $u$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.

\begin{table}[h!]
	\centering
	\begin{tabular}[t]{ll|llll}
		\hline
		\multicolumn{2}{c|}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
		\cline{3-6}
		Fields & Similarity & 1 & 2 & 3 & 4\\
		\hline
		First and Last Name & Levenstein & 0 & (0, .25] & (.25, .5] & (5, .1]\\
		\hline
		Age and Occupation & Binary & Agree & Disagree &  & \\
		\hline
	\end{tabular}
\caption{Construction of comparison vectors for accuracy study with simulated datasets}
\label{Tab:sadinle_simulation_cutoffs}
\end{table}

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot2} 
\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017. For each level of overlap and each level of error, we have 100 paired sets of 500 records.}
\label{fig:sadinle_simulation}
\end{center}
\end{figure}

@@

Lastly, we review the performance of \texttt{fabl} and \texttt{BRL} using loss functions that allow partial estimates of the linkage parameter. By leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match. In this context, we are no longer focused on finding all true matches, but rather protecting against false matches. Thus, instead of recall, we use the \emph{negative predictive value} (NPV), defined as the proportion of non-links that are actual non-matches. Mathematically, $NPV = \sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j = n_1 + j)$/$\sum_{j=1}^{n_2} I(\hat{Z}_j = n_1 + j)$. We continue to use the precision, which is renamed the \emph{positive predictive value} in this context. Lastly, we also report that rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)$.

@@

\begin{figure}[h!]
\begin{center}
	\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial} 
	\caption{Negative predictive value (NPV) and positive predictive value (PPV) on simulated datasets}
	\label{fig:sadinle_simulation_partial}
\end{center}
\end{figure}

@@

@@

\begin{table}[h!]
	\centering
	\begin{tabular}{rrr}
		\hline
		& m & u \\ 
		\hline
		fname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
		lname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
		day & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{30}, \frac{29}{30}\right)$ \\ 
		month & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
		year & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
		\hline
	\end{tabular}
\caption{Distributions used for $m$ and $u$ probabilties in simulation studies}\label{Tab:distributions}
\end{table}

\begin{figure}[h!]

@@

\begin{figure}[h!]

@@

@@

Specifically, we demonstrate how partitioning and hashing the data through \texttt{fabl} method allows us undertake significantly larger linkage tasks. We concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{} to create two larger files, each of 20,000 records (40,000 records total). Under standard Fellegi Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting a final comparison matrix about 6.4 GB in size. Under default settings on a personal computer, \texttt{R} does not allow storing an object this size, so \texttt{BRL} is not usable on this linkage task. Through \texttt{fabl} however, using just one machine, we sequentially compare records across chunks, hash results, and then synthesize summary statistics for the entire simulation.

For simplicity, we partition one dataset into 20 smaller chunks, and leave the second dataset fully intact. We compare records, hash results, and then synthesize summary statistics for all 20 chunk comparisons. The resulting data object is now only 90 MB, about 1\% the size of the object required under the standard method. Executed sequentially, these comparisons and the Gibbs sampler take about one hour to run. However, this could be substantially sped up using distributed computing.

Turning to evaluation metrics, this simulation achieved 96.5\% recall and 97.7\% precision, with an overall F-measure of 97.1\% F-measure. The reader will note that this slightly worse performance than witnessed in the smaller simulation studies; this is expected because it is naturally more difficult to link more files with the same amount of information. With more linkage fields \texttt{fabl} maintains the high accuracy seen above.

%\textcolor{blue}{This is with post processing already. All I meant here was that is that small tasks require less information. If you do a $500 \times 500$ linkage task and you only have first and last name (for example), thats likely enough to get all the correct matches. However, once you do $20000 \times 20000$, its more likely that more than one person in the second file has the same first and last name, so you need more information to distinguish. I mentioned before that the simulated files actually have 6 fields of information, but I only used 4 originally because thats what Sadinle did. I can use all 6 though to see what happens?}

\section{Application to Civilian Casualties from the El Salvadoran Civil War}
\label{sec:case-studies}

@@

{
\color{blue} For Beka: You wrote that you wanted more detail here, and I added some of it (with some citations). I believe I have all the important informatoin, just more concise. If I included Sadinle's level of detail, I worried about plagarism.
\vspace{1em}
}

The country of El Salvador was immersed in civil war from 1980 to 1991,
and throughout the time, several organizations attempted to document
casualties of the conflict. When estimating the total number of
casualties, one cannot simply sum the numbers recorded by each
organization, as it is likely that the same individuals are recorded in
multiple casualty lists. This is an important step in estimating the total number of deaths from a conflict using multiple systems estimation \citep{lum2013applications} Thanks to the Human Rights Data Analysis Group (HRDAG), we utilize 
two different sources, namely,  El Rescate - Tutela
Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
acronym in Spanish). The ERTL dataset consists of digitized denunciations that
had been published throughout the conflict, and additional investigation was required before the ERTL recorded them as human rights abuses, giving us confidence in its accuracy \cite{howland2008rescate}. The CDHES dataset consists
of casualties that had been reported directly to the organization, and
later digitized; since reports were taken soon after the events occured, this dataset is also thought to be fairly reliable. \cite{ball2000salvadoran}.

@@

In this paper, we utilize records that have non-missing entries for given and last name, which results 
in \(n_1 = 4420\) files in CHDES and \(n_2 = 1323\) files
in ERTL. We standardized names to account for common misspellings
in the Spanish language. The string distance used is a modified Levenstein
distance to account for the fact that second names are often omitted.
Place of death is recorded by municipality and department within that
municipality; however, since department was missing in 95\% of records
in CHDES and 80\% of records in ERTL, this was excluded from our analyses. 
Thus, we conduct record linkage using given name, last name,
municipality, and day, month, and year of death. We again use flat
priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters. To our knowledge, we have followed the same guidelines/settings in \cite{sadinle_bayesian_2017}.
\begin{table}
\begin{tabular}[h!]{ll|llll}
	\hline
	\multicolumn{2}{c|}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
	\cline{3-6}
	Fields & Similarity & 1 & 2 & 3 & 4\\
	\hline
	First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, .5] & (5, 1]\\
	\hline
	Year of Death & Absolute Difference & 0 & 1 & 2 & 3+\\
	\hline
	Month of Death & Absolute Difference & 0 & 1 & 2-3 & 4+\\
	\hline
	Day of Death & Absolute Difference & 0 & 1-2 & 3-7 & 8+\\
	\hline
	Municipality & Binary & Agree & Disagree &  & \\
	\hline
\end{tabular}
\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \citep{sadinle2018bayesian}. This set up leads to 2048 possible agreement patterns in total}\label{Tab:el_salvador_cutoffs_1}
\end{table}
